---
title: "Bellabeat"
author: "Pilar Wilches"
date: "2023-10-03"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

## Analysis Process for the case study: Bellabeat

### Goal: Analyse data from smart devices to understand consumer use of smart device

First of all we install packages and libraries for data clean and
visualization

```{r import packages and libraries}
options(repos = "https://cran.rstudio.com/")

install.packages("here")
install.packages("skimr")
install.packages("janitor")
install.packages("ggplot2")

library("here")
library("skimr")
library("janitor")
library("dplyr")
library("lubridate")
library("readr")
library("ggplot2")
```

# Work with file dailyIntensities_merged.csv

## 1. Import Data and previsualization

Now, I import the file: dailyIntensities_merged.csv and create the data
frame, With this data frame I want know the relationship between the
days of the week and the person activity.

```{r upload data Intensities}
setwd("/xxx/xxxx/Bellabeat") #change the path here to the path where your file is located
dailyIntensities_df <- read_csv("dailyIntensities_merged.csv")
```

Then, I preview the data with different functions for review cols, rows,
type of data.

```{r preview of data dailyIntensities1}

View(dailyIntensities_df)

glimpse(dailyIntensities_df)
```

```{r preview of data dailyIntensities2}

colnames(dailyIntensities_df)
```

## 2. Data cleaning

I check the consistence of data frame

```{r check columns dailyIntensities}
clean_names(dailyIntensities_df)
```

When I apply this function, I see that the name of the columns can be
improved according to the syntax of clean code, so I change the column
names type snake case

```{r rename colums to snake_case dailyIntensities}

 rename_columns_to_snake_case <- function(dataframe){
  dataframe %>% 
    rename_with(~tolower(gsub("([a-z0-9])([A-Z])", "\\1_\\2", .)),.cols = everything())
 }

dailyIntensities_new_df <- rename_columns_to_snake_case(dailyIntensities_df)

View(dailyIntensities_new_df)
```

## 3. Manipulation/Processing of data

As I want see the day of week has more or less activity then I work with
the field activity_data and I convert this field char in date. For this
I use the library lubridate.

```{r create column day_of_week dailyIntensities}
dailyIntensities_new_df <- dailyIntensities_new_df %>% 
  mutate(day_of_week = weekdays(as.Date(activity_day, format="%m/%d/%Y")) )
```

Now, it is very important to have statistics, below is the comparison:

Day of week vrs. Average active minutes

```{r statistics mean_active_minutes dailyIntensities}
mean_active_minutes_df <- dailyIntensities_new_df %>% 
  group_by(day_of_week) %>% 
  summarize(mean_active_minutes =      round(mean(very_active_minutes+fairly_active_minutes+lightly_active_minutes))) %>% 
  arrange(mean_active_minutes)

```

Day of week vrs. Sedentary minutes

```{r mean_sedentary_minutes dailyIntensities}
dailyIntensities_new_df %>% 
  group_by(day_of_week) %>% 
  summarize(mean_sedentary_minutes =round(mean(sedentary_minutes))) %>% 
  arrange(desc(mean_sedentary_minutes))
```

# Work with file dailyActivity_merged.csv

## 1. Import Data and previsualization

Now, I import the file: dailyActivity_merged.csv and create the data
frame, With this data frame I want know the relationship between the
days of the week and the calories, steps and activity in general.

```{r upload data activity}
setwd("/Users/maria/Bellabeat")
dailyActivity_df <- read_csv("dailyActivity_merged.csv")
```

Then, I preview the data with different functions for review cols, rows,
type of data.

```{r preview of data dailyActivity1}

View(dailyActivity_df)

glimpse(dailyActivity_df)
```

This visualization is very important, for instance, the column
ActivityDay is character and must be date type for the analysis.

```{r preview of data dailyActivity2}

colnames(dailyActivity_df)
```

## 2. Data cleaning

I check the consistence of data frame

```{r check columns dailyActivity}
clean_names(dailyActivity_df)
```

When I apply this function, I see that the name of the columns can be
improved according the syntax of clean code, so I apply the function
created previously, which changes the column names like snake case

```{r rename colums to snake_case dailyActivity}

dailyActivity_new_df <- rename_columns_to_snake_case(dailyActivity_df)

View(dailyActivity_new_df)
```

## 3. Manipulation/Processing of data

And this point, I used the same code for change and add the column name
day_of_week

```{r create column day_of_week dailyActivity}
dailyActivity_new_df <- dailyActivity_new_df %>% 
  mutate(day_of_week = weekdays(as.Date(activity_date, format="%m/%d/%Y")) )
```

Check the new column:

```{r check add column dailyActivity}
 View(dailyActivity_new_df)
```

Now, it's statistics' time:

Day of week vrs. Calories

```{r statistics mean_calories dailyActivity}
 dailyActivity_new_df %>% 
  group_by(day_of_week) %>% 
  summarize(mean_calories = round(mean(calories,na.rm = TRUE)), 
            mean_steps = round(mean(total_steps,na.rm = TRUE))) %>% 
  arrange(desc(mean_calories))

 mean_calories_df <- dailyActivity_new_df %>% 
  group_by(day_of_week) %>% 
  summarize(mean_calories = round(mean(calories, na.rm = TRUE))) %>% 
  arrange(desc(mean_calories))
 
  mean_total_steps_df <- dailyActivity_new_df %>% 
  group_by(day_of_week) %>% 
  summarize(mean_steps = round(mean(total_steps, na.rm = TRUE))) %>% 
  arrange(desc(mean_steps))

 
```

# Visualization of data

In this step, I'll show the different graphics of my data:

ggplot

```{r graphic minutes very active}
ggplot(data = mean_active_minutes_df) +
  geom_col(mapping = aes(x= day_of_week, y=mean_active_minutes), fill="#2e8540")+
  labs(title = "Day of Week vrs. Minutes Very Active",
       subtitle = "Check out which days are more intense",
       caption = "FitBit Fitness Tracker Data")

```

```{r graphic calories}
ggplot(data = mean_calories_df) +
  geom_col(mapping = aes(x= day_of_week, y=mean_calories), fill="#205493")+
  labs(title = "Day of Week vrs. Calories",
       subtitle = "What days have more calories burned?",
       caption = "FitBit Fitness Tracker Data")

```

```{r calories and distance, message=FALSE, warning=FALSE}
ggplot(data = dailyActivity_new_df) +
  geom_point(mapping = aes(x= total_distance, y=calories), color="#e31c3d") +
  geom_smooth(mapping = aes(x= total_distance, y=calories)) +
  labs(title = "Distance traveled vrs. Calories burned",
       subtitle = "Let's find out what registered device",
       caption = "FitBit Fitness Tracker Data")
```

```{r sedentary vrs total steps}

ggplot(data=dailyActivity_new_df, aes(x=total_steps, y=sedentary_minutes)) + 
  geom_point() +
  labs(title = "Sedentary vrs. Total Steps ",
       caption = "FitBit Fitness Tracker Data")
  
```

# Technical Notes

-   I could have named all variables with clean code I was able to used
    a function to change the data to the day of week and not repeat
    code.

-   As I progressed in the project, I became more familiar with the
    files, so I realized note that I could have started with the
    activity file and not Intensitive file csv.

-   In data cleaning I didn't delete the day with zero activity because
    I don't know if it is a mistake, this situation in a real case, I
    would ask the organization.

# Conclusions

-   Develop sporting events on Saturdays. This day presents the highest
    intensity of movement reported in the smart devices. The figures
    reveal that on average on Saturday, women have 240 minutes of high
    physical activity (4 hours)

-   Thursdays are the day on average that the fewest calories are
    burned. But how do users know if they are burning enough calories to
    know if they have healthy habits? The strategy I propose is that on    Thursdays,  through  email and the company's social media accounts, to show content on healthy habits in accordance with what is suggested by the OMS.

-   Create an awareness campaign, making visible the days of greatest
    sedentary lifestyle.
